### 2019-11-01 submission/manuscript.tex
Respond to reviewers' comments

Proton pump inhibitor (PPI) use has been associated with microbiota
alterations and susceptibility to \emph{Clostridium difficile}
infections (CDIs) in humans. We assessed how PPI treatment alters the
fecal microbiota and whether treatment promotes CDIs in a mouse model.
Mice receiving a PPI treatment were gavaged with 40 mg/kg of omeprazole
during a 7-day pretreatment phase, the day of \emph{C. difficile}
challenge, and the following 9 days. We found that mice treated with
omeprazole were not colonized by \emph{C. difficile}. When omeprazole
treatment was combined with a single clindamycin treatment, one cage of
mice remained resistant to \emph{C. difficile} colonization, while the
other cage was colonized. Treating mice with only clindamycin followed
by challenge resulted in \emph{C. difficile} colonization. 16S rRNA gene
sequencing analysis revealed that omeprazole had minimal impact on the
structure of the murine microbiota throughout the 16 days of omeprazole
exposure. These results suggest omeprazole treatment alone is not
sufficient to disrupt microbiota resistance to \emph{C. difficile}
infection in mice that are normally resistant in the absence of
antibiotic treatment.

@@

Antibiotics are the primary risk factor for \emph{Clostridium difficile}
infections (CDIs), but other factors may also increase a person's risk.
In epidemiological studies, proton pump inhibitor (PPI) use has been
associated with CDI incidence and recurrence. PPIs have also been
associated with alterations in the human intestinal microbiota in
observational and interventional studies. We evaluated the effects of
the PPI omeprazole on the structure of the murine intestinal microbiota
and its ability to disrupt colonization resistance to \emph{C.
difficile}. We found omeprazole treatment had minimal impact on the
murine fecal microbiota and did not promote \emph{C. difficile}
colonization. Further studies are needed to determine whether other
factors such as the composition of the starting bacterial community,
comorbidities, and use of additional medications contribute to the
association between PPIs and CDIs seen in humans or whether aspects of
murine physiology may limit its utility to test these types of
hypotheses.

@@

Antibiotics have a large impact on the intestinal microbiome and are a
primary risk factor for developing \emph{Clostridium difficile}
infections (CDIs) (1). It is less clear whether other human medications
that impact the microbiota also influence \emph{C. difficile}
colonization resistance. Multiple epidemiological studies have suggested
an association between proton pump inhibitor (PPI) use and incidence or
recurrence of CDIs (2--5). There have also been a number of large cohort
studies and interventional clinical trials that demonstrated specific
alterations in the intestinal microbiome were associated with PPI use
(4, 6). PPI-associated microbiota changes have been attributed to the
ability of PPIs to increase stomach acid pH which may promote the
survival of oral and pathogenic bacteria (4, 6). In human fecal samples,
PPI use results in increases in \emph{Enterococcaceae},
\emph{Lactobacillaceae}, \emph{Micrococcaceae}, \emph{Staphylococcaceae}
and \emph{Streptococcaceae} and decreases in \emph{Ruminococcaceae}
(6--9). Several of these taxa have also been associated with \emph{C.
difficile} colonization in humans (10).\\
Unfortunately, the studies suggesting a link between PPIs and \emph{C.
difficile} were retrospective and did not evaluate changes in the
microbiome (2, 3, 5). Thus, it is unclear whether the gastrointestinal
microbiome changes associated with PPI use explain the association
between PPIs and CDIs. Additionally, epidemiological studies have a
limited capacity to address potential confounders and comorbidities in
patients that were on PPIs and developed CDIs or recurrent CDIs (2, 5).
Here, we evaluated the impact of daily PPI treatment with omeprazole on
the murine microbiome and susceptibility to \emph{C. difficile}
colonization in relation to clindamycin, an antibiotic that perturbs the
microbiome enough to allow \emph{C. difficile} to colonize but is mild
enough that \emph{C. difficile} is cleared within 10 days (11).

@@

abundance in our mice. Although our fecal microbiota findings are

@@

treatment (13--15), there were insufficient controls to attribute the

@@

with antibiotics (13, 14). The other mouse study demonstrated

@@

treated with just esomeprazole for comparison (15). We tested the same

@@

colonization without antibiotic treatment (16), ensuring there was not

@@

In contrast to previous work with PPIs (13--15), omeprazole did not

@@

mouse model. Gastrointestinal physiological differences, particularly
the higher stomach pH in mice (pH 3-4) compared to humans (pH 1) (17)
could also explain why omeprazole had a limited impact on the murine
microbiome. The microbiota and physiological differences between humans
and mice may limit the usefulness of employing mouse models to study the
impact of PPIs on the microbiota and CDIs.\\

@@

recurrence. This study addressed the impact of PPIs with or without
antibiotics on a murine model of CDI, and found PPIs did not promote
\emph{C. difficile} colonization. The epidemiological evidence linking
PPIs to CDIs is primarily from observational studies, which makes
determining causality and whether other risk factors play a role
challenging (18). Future studies are needed to determine whether age,
other comorbidities and bacterial strains that are less common in mice
can increase the risk of CDIs or recurrent CDIs when combined with PPI
treatment.

@@

\textbf{Drug treatments.} Omeprazole (Sigma Aldrich) was prepared in a
vehicle solution of 40\% polyethylene glycol 400 (Sigma-Aldrich) in
phosphate buffered saline. Omeprazole was prepared from 20 mg/mL frozen
aliquots and diluted to an 8 mg/mL prior to gavage. All mice received 40
mg/kg omeprazole (a dose previously used in mouse experiments (15)) or
vehicle solution once per day through the duration of the experiment
with treatment starting 7 days before \emph{C. difficile} challenge
(Figure 1A). One day prior to \emph{C. difficile} challenge, 2 groups of
mice received an intraperitoneal injection of 10 mg/kg clindamycin or
sterile saline vehicle (11). All drugs were filter sterilized through a
0.22 micron syringe filter before administration to animals.

@@

\textbf{16S rRNA gene sequencing.} DNA for 16S rRNA gene sequencing was
extracted from 10-50 mg fecal pellet from each mouse using the DNeasy
Powersoil HTP 96 Kit (Qiagen) and an EpMotion 5075 automated pipetting
system (Eppendorf). The 16S rRNA sequencing library was prepared as
described previously (19). In brief, the ZymoBIOMICS\textsuperscript{TM}
Microbial Community DNA Standard (Zymo, CA, USA) was used as a mock
community (20) and water was used as a negative control. The V4
hypervariable region of the 16S rRNA gene was amplified with Accuprime
Pfx DNA polymerase (Thermo Fisher Scientific) using previously described
custom barcoded primers (19). The 16S rRNA amplicon library was
sequenced with the MiSeq (Illumina). Amplicons were cleaned up and
normalized with the SequalPrep Normalization Plate Kit (ThermoFisher
Scientific) and pooled amplicons were quantified with the KAPA library
quantification kit (KAPA Biosystems).

\textbf{16S rRNA gene sequence analysis.} mothur (v1.40.5) was used for
all sequence processing steps (21) using a previously published protocol
(19). In brief, forward and reverse reads for each sample were combined
and low-quality sequences and chimeras were removed. Duplicate sequences
were merged, before taxonomy assignment using a modified version (v16)
of the Ribosomal Database Project reference database (v11.5) with an
80\% cutoff. Operational taxonomic units (OTUs) were assigned with the
opticlust clustering algorithm using a 97\% similarity threshold. To
adjust for uneven sequencing across samples, all samples were rarefied
to 3,000 sequences, 1,000 times. PCoAs were generated based on
Bray-Curtis distance. R (v.3.5.1) was used to generate figures and
perform statistical analysis.

@@

\hypertarget{refs}{}
\hypertarget{ref-Schubert2015}{}
1. \textbf{Schubert AM}, \textbf{Sinani H}, \textbf{Schloss PD}. 2015.
Antibiotic-induced alterations of the murine gut microbiota and
subsequent effects on colonization resistance against clostridium
difficile. mBio \textbf{6}.
doi:\href{https://doi.org/10.1128/mbio.00974-15}{10.1128/mbio.00974-15}.

\hypertarget{ref-tariq2017association}{}
2. \textbf{Tariq R}, \textbf{Singh S}, \textbf{Gupta A}, \textbf{Pardi
DS}, \textbf{Khanna S}. 2017. Association of gastric acid suppression
with recurrent clostridium difficile infection: A systematic review and
meta-analysis. JAMA internal medicine \textbf{177}:784--791.

\hypertarget{ref-nehra2018proton}{}
3. \textbf{Nehra AK}, \textbf{Alexander JA}, \textbf{Loftus CG},
\textbf{Nehra V}. 2018. Proton pump inhibitors: Review of emerging
concerns, pp. 240--246. \emph{In} Mayo clinic proceedings. Elsevier.

\hypertarget{ref-Naito2018}{}
4. \textbf{Naito Y}, \textbf{Kashiwagi K}, \textbf{Takagi T},
\textbf{Andoh A}, \textbf{Inoue R}. 2018. Intestinal dysbiosis secondary
to proton-pump inhibitor use. Digestion \textbf{97}:195--204.
doi:\href{https://doi.org/10.1159/000481813}{10.1159/000481813}.

\hypertarget{ref-Elias2019}{}
5. \textbf{Elias E}, \textbf{Targownik LE}. 2019. The clinician's guide
to proton pump inhibitor related adverse events. Drugs
\textbf{79}:715--731.
doi:\href{https://doi.org/10.1007/s40265-019-01110-3}{10.1007/s40265-019-01110-3}.

\hypertarget{ref-Imhann2017}{}
6. \textbf{Imhann F}, \textbf{Vila AV}, \textbf{Bonder MJ},
\textbf{Manosalva AGL}, \textbf{Koonen DP}, \textbf{Fu J},
\textbf{Wijmenga C}, \textbf{Zhernakova A}, \textbf{Weersma RK}. 2017.
The influence of proton pump inhibitors and other commonly used
medication on the gut microbiota. Gut Microbes \textbf{8}:351--358.
doi:\href{https://doi.org/10.1080/19490976.2017.1284732}{10.1080/19490976.2017.1284732}.

\hypertarget{ref-Imhann2015}{}
7. \textbf{Imhann F}, \textbf{Bonder MJ}, \textbf{Vila AV}, \textbf{Fu
J}, \textbf{Mujagic Z}, \textbf{Vork L}, \textbf{Tigchelaar EF},
\textbf{Jankipersadsing SA}, \textbf{Cenit MC}, \textbf{Harmsen HJM},
\textbf{Dijkstra G}, \textbf{Franke L}, \textbf{Xavier RJ},
\textbf{Jonkers D}, \textbf{Wijmenga C}, \textbf{Weersma RK},
\textbf{Zhernakova A}. 2015. Proton pump inhibitors affect the gut
microbiome. Gut \textbf{65}:740--748.
doi:\href{https://doi.org/10.1136/gutjnl-2015-310376}{10.1136/gutjnl-2015-310376}.

\hypertarget{ref-Freedberg2015}{}
8. \textbf{Freedberg DE}, \textbf{Toussaint NC}, \textbf{Chen SP},
\textbf{Ratner AJ}, \textbf{Whittier S}, \textbf{Wang TC}, \textbf{Wang
HH}, \textbf{Abrams JA}. 2015. Proton pump inhibitors alter specific
taxa in the human gastrointestinal microbiome: A crossover trial.
Gastroenterology \textbf{149}:883--885.e9.
doi:\href{https://doi.org/10.1053/j.gastro.2015.06.043}{10.1053/j.gastro.2015.06.043}.

\hypertarget{ref-maier2018extensive}{}
9. \textbf{Maier L}, \textbf{Pruteanu M}, \textbf{Kuhn M},
\textbf{Zeller G}, \textbf{Telzerow A}, \textbf{Anderson EE},
\textbf{Brochado AR}, \textbf{Fernandez KC}, \textbf{Dose H},
\textbf{Mori H}, \textbf{others}. 2018. Extensive impact of
non-antibiotic drugs on human gut bacteria. Nature \textbf{555}:623.

\hypertarget{ref-Schuberte01021-14}{}
10. \textbf{Schubert AM}, \textbf{Rogers MAM}, \textbf{Ring C},
\textbf{Mogle J}, \textbf{Petrosino JP}, \textbf{Young VB},
\textbf{Aronoff DM}, \textbf{Schloss PD}. 2014. Microbiome data
distinguish patients with clostridium difficile infection and non-c.
difficile-associated diarrhea from healthy controls. mBio \textbf{5}.
doi:\href{https://doi.org/10.1128/mBio.01021-14}{10.1128/mBio.01021-14}.

\hypertarget{ref-Jenior2018}{}
11. \textbf{Jenior ML}, \textbf{Leslie JL}, \textbf{Young VB},
\textbf{Schloss PD}. 2018. Clostridium difficile alters the structure
and metabolism of distinct cecal microbiomes during initial infection to
promote sustained colonization. mSphere \textbf{3}.
doi:\href{https://doi.org/10.1128/msphere.00261-18}{10.1128/msphere.00261-18}.

\hypertarget{ref-Yasutomi2018}{}
12. \textbf{Yasutomi E}, \textbf{Hoshi N}, \textbf{Adachi S},
\textbf{Otsuka T}, \textbf{Kong L}, \textbf{Ku Y}, \textbf{Yamairi H},
\textbf{Inoue J}, \textbf{Ishida T}, \textbf{Watanabe D}, \textbf{Ooi
M}, \textbf{Yoshida M}, \textbf{Tsukimi T}, \textbf{Fukuda S},
\textbf{Azuma T}. 2018. Proton pump inhibitors increase the
susceptibility of mice to oral infection with enteropathogenic bacteria.
Digestive Diseases and Sciences \textbf{63}:881--889.
doi:\href{https://doi.org/10.1007/s10620-017-4905-3}{10.1007/s10620-017-4905-3}.

\hypertarget{ref-Kaur2007}{}
13. \textbf{Kaur S}, \textbf{Vaishnavi C}, \textbf{Prasad KK},
\textbf{Ray P}, \textbf{Kochhar R}. 2007. Comparative role of antibiotic
and proton pump inhibitor in ExperimentalClostridium difficileInfection
in mice. Microbiology and Immunology \textbf{51}:1209--1214.
doi:\href{https://doi.org/10.1111/j.1348-0421.2007.tb04016.x}{10.1111/j.1348-0421.2007.tb04016.x}.

\hypertarget{ref-kaur2011effect}{}
14. \textbf{Kaur S}, \textbf{Vaishnavi C}, \textbf{Prasad KK},
\textbf{Ray P}, \textbf{Kochhar R}. 2011. Effect of lactobacillus
acidophilus \& epidermal growth factor on experimentally induced
clostridium difficile infection. The Indian journal of medical research
\textbf{133}:434.

\hypertarget{ref-hung2015proton}{}
15. \textbf{Hung Y-P}, \textbf{Ko W-C}, \textbf{Chou P-H}, \textbf{Chen
Y-H}, \textbf{Lin H-J}, \textbf{Liu Y-H}, \textbf{Tsai H-W}, \textbf{Lee
J-C}, \textbf{Tsai P-J}. 2015. Proton-pump inhibitor exposure aggravates
clostridium difficile--associated colitis: Evidence from a mouse model.
The Journal of infectious diseases \textbf{212}:654--663.

\hypertarget{ref-Jenior2017}{}
16. \textbf{Jenior ML}, \textbf{Leslie JL}, \textbf{Young VB},
\textbf{Schloss PD}. 2017. Clostridium difficile colonizes alternative
nutrient niches during infection across distinct murine gut microbiomes.
mSystems \textbf{2}.
doi:\href{https://doi.org/10.1128/msystems.00063-17}{10.1128/msystems.00063-17}.

\hypertarget{ref-Hugenholtz2017}{}
17. \textbf{Hugenholtz F}, \textbf{Vos WM de}. 2017. Mouse models for
human intestinal microbiota research: A critical evaluation. Cellular
and Molecular Life Sciences \textbf{75}:149--160.
doi:\href{https://doi.org/10.1007/s00018-017-2693-8}{10.1007/s00018-017-2693-8}.

\hypertarget{ref-Eze2017}{}
18. \textbf{Eze P}, \textbf{Balsells E}, \textbf{Kyaw MH}, \textbf{Nair
H}. 2017. Risk factors for clostridium difficile infections an overview
of the evidence base and challenges in data synthesis. Journal of Global
Health \textbf{7}.
doi:\href{https://doi.org/10.7189/jogh.07.010417}{10.7189/jogh.07.010417}.

\hypertarget{ref-Kozich2013}{}
19. \textbf{Kozich JJ}, \textbf{Westcott SL}, \textbf{Baxter NT},
\textbf{Highlander SK}, \textbf{Schloss PD}. 2013. Development of a
dual-index sequencing strategy and curation pipeline for analyzing
amplicon sequence data on the MiSeq illumina sequencing platform.
Applied and Environmental Microbiology \textbf{79}:5112--5120.
doi:\href{https://doi.org/10.1128/aem.01043-13}{10.1128/aem.01043-13}.

\hypertarget{ref-Sze2019}{}
20. \textbf{Sze MA}, \textbf{Schloss PD}. 2019. The impact of DNA
polymerase and number of rounds of amplification in PCR on 16S rRNA gene
sequence data. mSphere \textbf{4}.
doi:\href{https://doi.org/10.1128/msphere.00163-19}{10.1128/msphere.00163-19}.

\hypertarget{ref-Schloss2009}{}
21. \textbf{Schloss PD}, \textbf{Westcott SL}, \textbf{Ryabin T},
\textbf{Hall JR}, \textbf{Hartmann M}, \textbf{Hollister EB},
\textbf{Lesniewski RA}, \textbf{Oakley BB}, \textbf{Parks DH},
\textbf{Robinson CJ}, \textbf{Sahl JW}, \textbf{Stres B},
\textbf{Thallinger GG}, \textbf{Horn DJV}, \textbf{Weber CF}. 2009.
Introducing mothur: Open-source, platform-independent,
community-supported software for describing and comparing microbial
communities. Applied and Environmental Microbiology
\textbf{75}:7537--7541.
doi:\href{https://doi.org/10.1128/aem.01541-09}{10.1128/aem.01541-09}.

>>>>

Proton pump inhibitor (PPI) use has been associated with microbiota
alterations and susceptibility to \emph{Clostridium}
(\emph{Clostridioides}) \emph{difficile} infections (CDIs) in humans. We
assessed how PPI treatment alters the fecal microbiota and whether
treatment promotes CDIs in a mouse model. Mice receiving a PPI treatment
were gavaged with 40 mg/kg of omeprazole during a 7-day pretreatment
phase, the day of \emph{C. difficile} challenge, and the following 9
days. We found that mice treated with omeprazole were not colonized by
\emph{C. difficile}. When omeprazole treatment was combined with a
single clindamycin treatment, one cage of mice remained resistant to
\emph{C. difficile} colonization, while the other cage was colonized.
Treating mice with only clindamycin followed by challenge resulted in
\emph{C. difficile} colonization. 16S rRNA gene sequencing analysis
revealed that omeprazole had minimal impact on the structure of the
murine microbiota throughout the 16 days of omeprazole exposure. These
results suggest omeprazole treatment alone is not sufficient to disrupt
microbiota resistance to \emph{C. difficile} infection in mice that are
normally resistant in the absence of antibiotic treatment.

@@

Antibiotics are the primary risk factor for \emph{Clostridium}
(\emph{Clostridioides}) \emph{difficile} infections (CDIs), but other
factors may also increase a person's risk. In epidemiological studies,
proton pump inhibitor (PPI) use has been associated with CDI incidence
and recurrence. PPIs have also been associated with alterations in the
human intestinal microbiota in observational and interventional studies.
We evaluated the effects of the PPI omeprazole on the structure of the
murine intestinal microbiota and its ability to disrupt colonization
resistance to \emph{C. difficile}. We found omeprazole treatment had
minimal impact on the murine fecal microbiota and did not promote
\emph{C. difficile} colonization. Further studies are needed to
determine whether other factors contribute to the association between
PPIs and CDIs seen in humans or whether aspects of murine physiology may
limit its utility to test these types of hypotheses.

@@

Antibiotics have a large impact on the intestinal microbiome and are a
primary risk factor for developing \emph{Clostridium}
(\emph{Clostridioides}) \emph{difficile} infections (CDIs) (1). It is
less clear whether other human medications that impact the microbiota
also influence \emph{C. difficile} colonization resistance. Multiple
epidemiological studies have suggested an association between proton
pump inhibitor (PPI) use and incidence or recurrence of CDIs (2--5).
There have also been a number of large cohort studies and interventional
clinical trials that demonstrated specific alterations in the intestinal
microbiome were associated with PPI use (4, 6). PPI-associated
microbiota changes have been attributed to the ability of PPIs to
increase stomach acid pH which may promote the survival of oral and
pathogenic bacteria (4, 6). In human fecal samples, PPI use results in
increases in \emph{Enterococcaceae}, \emph{Lactobacillaceae},
\emph{Micrococcaceae}, \emph{Staphylococcaceae} and
\emph{Streptococcaceae} and decreases in \emph{Ruminococcaceae} (6--9).
Several of these taxa have also been associated with \emph{C. difficile}
colonization in humans (10).\\
Unfortunately, the studies suggesting a link between PPIs and \emph{C.
difficile} were retrospective and did not evaluate changes in the
microbiome (2, 3, 5). Thus, it is unclear whether the gastrointestinal
microbiome changes associated with PPI use explain the association
between PPIs and CDIs. Additionally, epidemiological studies have a
limited capacity to address potential confounders and comorbidities in
patients that were on PPIs and developed CDIs or recurrent CDIs (2, 5).
Here, we evaluated the impact of daily PPI treatment with omeprazole on
the murine microbiome and susceptibility to \emph{C. difficile}
colonization in relation to clindamycin, an antibiotic that perturbs the
microbiome enough to allow \emph{C. difficile} to colonize but is mild
enough that \emph{C. difficile} is cleared within 10 days (11).

@@

abundance in our mice. Interestingly, other groups examining fecal
microbiota communities before and after PPI administration to healthy
cats and infants with gastroesophageal reflux disease, found PPIs have
minimal effects on fecal bacterial community structures, although there
were a few significant changes in specific genera (13, 14). One
limitation of our study is that there were only 4-5 mice per group,
which may have limited our ability to identify PPI-induced changes in
specific bacteria genera. Although our fecal microbiota findings are

@@

treatment (15--17), there were insufficient controls to attribute the

@@

with antibiotics (15, 16). The other mouse study demonstrated

@@

treated with just esomeprazole for comparison (17). We tested the same

@@

colonization without antibiotic treatment (18), ensuring there was not

@@

In contrast to previous work with PPIs (15--17), omeprazole did not

@@

mouse model. While the intestinal microbiomes of both humans and mice
are dominated by the \emph{Bacteroidetes} and \emph{Firmicutes} phyla,
there are significant differences in the relative abundances of genera
that are present and some genera are unique to each mammal (19),
differences that may partly explain our results. Gastrointestinal
physiological differences, particularly the higher stomach pH in mice
(pH 3-4) compared to humans (pH 1) (19) could also explain why
omeprazole had a limited impact on the murine microbiome. The microbiota
and physiological differences between humans and mice may limit the
usefulness of employing mouse models to study the impact of PPIs on the
microbiota and CDIs.\\

@@

recurrence. The type of \emph{C. difficile} strain type could also be an
important contributing factor, however our study was limited in that we
only tested \emph{C. difficile} 630 (ribotype 012). This study addressed
the impact of PPIs with or without antibiotics on a murine model of CDI,
and found PPIs did not promote \emph{C. difficile} colonization. The
epidemiological evidence linking PPIs to CDIs is primarily from
observational studies, which makes determining causality and whether
other risk factors play a role challenging (20). Future studies are
needed to determine whether age, other comorbidities and bacterial
strains that are less common in mice can increase the risk of CDIs or
recurrent CDIs when combined with PPI treatment.

@@

\textbf{Drug treatments.} Omeprazole (Sigma Aldrich) was prepared in a
vehicle solution of 40\% polyethylene glycol 400 (Sigma-Aldrich) in
phosphate buffered saline. Omeprazole was prepared from 20 mg/mL frozen
aliquots and diluted to an 8 mg/mL prior to gavage. All mice received 40
mg/kg omeprazole (a dose previously used in mouse experiments (17)) or
vehicle solution once per day through the duration of the experiment
with treatment starting 7 days before \emph{C. difficile} challenge
(Figure 1A). Although the omeprazole dose administered to mice is higher
than the recommended dose for humans, omeprazole has a shorter half-life
in mice compared to humans (21) and lacks an enteric coating (22). One
day prior to \emph{C. difficile} challenge, 2 groups of mice received an
intraperitoneal injection of 10 mg/kg clindamycin or sterile saline
vehicle (11). All drugs were filter sterilized through a 0.22 micron
syringe filter before administration to animals.

@@

\textbf{16S rRNA gene sequencing.} DNA for 16S rRNA gene sequencing was
extracted from 10-50 mg fecal pellet from each mouse using the DNeasy
Powersoil HTP 96 Kit (Qiagen) and an EpMotion 5075 automated pipetting
system (Eppendorf). The 16S rRNA sequencing library was prepared as
described previously (23). In brief, the ZymoBIOMICS\textsuperscript{TM}
Microbial Community DNA Standard (Zymo, CA, USA) was used as a mock
community (24) and water was used as a negative control. The V4
hypervariable region of the 16S rRNA gene was amplified with Accuprime
Pfx DNA polymerase (Thermo Fisher Scientific) using previously described
custom barcoded primers (23). The 16S rRNA amplicon library was
sequenced with the MiSeq (Illumina). Amplicons were cleaned up and
normalized with the SequalPrep Normalization Plate Kit (ThermoFisher
Scientific) and pooled amplicons were quantified with the KAPA library
quantification kit (KAPA Biosystems).

\textbf{16S rRNA gene sequence analysis.} mothur (v1.40.5) was used for
all sequence processing steps (25) using a previously published protocol
(23). In brief, forward and reverse reads for each sample were combined
and low-quality sequences and chimeras were removed. Duplicate sequences
were merged, before taxonomy assignment using a modified version (v16)
of the Ribosomal Database Project reference database (v11.5) with an
80\% cutoff. Operational taxonomic units (OTUs) were assigned with the
opticlust clustering algorithm using a 97\% similarity threshold. To
adjust for uneven sequencing across samples, all samples were rarefied
to 3,000 sequences, 1,000 times. PCoAs were generated based on
Bray-Curtis distance. R (v.3.5.1) was used to generate figures and
perform statistical analysis.

@@

\hypertarget{refs}{}
\hypertarget{ref-Schubert2015}{}
1. Schubert AM, Sinani H, Schloss PD. 2015. Antibiotic-induced
alterations of the murine gut microbiota and subsequent effects on
colonization resistance against \emph{Clostridium difficile}. mBio 6.

\hypertarget{ref-tariq2017association}{}
2. Tariq R, Singh S, Gupta A, Pardi DS, Khanna S. 2017. Association of
gastric acid suppression with recurrent \emph{Clostridium difficile}
infection: A systematic review and meta-analysis. JAMA internal medicine
177:784--791.

\hypertarget{ref-nehra2018proton}{}
3. Nehra AK, Alexander JA, Loftus CG, Nehra V. 2018. Proton pump
inhibitors: Review of emerging concerns, pp. 240--246. \emph{In} Mayo
clinic proceedings. Elsevier.

\hypertarget{ref-Naito2018}{}
4. Naito Y, Kashiwagi K, Takagi T, Andoh A, Inoue R. 2018. Intestinal
dysbiosis secondary to proton-pump inhibitor use. Digestion 97:195--204.

\hypertarget{ref-Elias2019}{}
5. Elias E, Targownik LE. 2019. The clinician's guide to proton pump
inhibitor related adverse events. Drugs 79:715--731.

\hypertarget{ref-Imhann2017}{}
6. Imhann F, Vila AV, Bonder MJ, Manosalva AGL, Koonen DP, Fu J,
Wijmenga C, Zhernakova A, Weersma RK. 2017. The influence of proton pump
inhibitors and other commonly used medication on the gut microbiota. Gut
Microbes 8:351--358.

\hypertarget{ref-Imhann2015}{}
7. Imhann F, Bonder MJ, Vila AV, Fu J, Mujagic Z, Vork L, Tigchelaar EF,
Jankipersadsing SA, Cenit MC, Harmsen HJM, Dijkstra G, Franke L, Xavier
RJ, Jonkers D, Wijmenga C, Weersma RK, Zhernakova A. 2015. Proton pump
inhibitors affect the gut microbiome. Gut 65:740--748.

\hypertarget{ref-Freedberg2015}{}
8. Freedberg DE, Toussaint NC, Chen SP, Ratner AJ, Whittier S, Wang TC,
Wang HH, Abrams JA. 2015. Proton pump inhibitors alter specific taxa in
the human gastrointestinal microbiome: A crossover trial.
Gastroenterology 149:883--885.e9.

\hypertarget{ref-maier2018extensive}{}
9. Maier L, Pruteanu M, Kuhn M, Zeller G, Telzerow A, Anderson EE,
Brochado AR, Fernandez KC, Dose H, Mori H, others. 2018. Extensive
impact of non-antibiotic drugs on human gut bacteria. Nature 555:623.

\hypertarget{ref-Schuberte01021-14}{}
10. Schubert AM, Rogers MAM, Ring C, Mogle J, Petrosino JP, Young VB,
Aronoff DM, Schloss PD. 2014. Microbiome data distinguish patients with
\emph{Clostridium difficile} infection and non-\emph{C.
difficile}-associated diarrhea from healthy controls. mBio 5.

\hypertarget{ref-Jenior2018}{}
11. Jenior ML, Leslie JL, Young VB, Schloss PD. 2018. \emph{Clostridium
difficile} alters the structure and metabolism of distinct cecal
microbiomes during initial infection to promote sustained colonization.
mSphere 3.

\hypertarget{ref-Yasutomi2018}{}
12. Yasutomi E, Hoshi N, Adachi S, Otsuka T, Kong L, Ku Y, Yamairi H,
Inoue J, Ishida T, Watanabe D, Ooi M, Yoshida M, Tsukimi T, Fukuda S,
Azuma T. 2018. Proton pump inhibitors increase the susceptibility of
mice to oral infection with enteropathogenic bacteria. Digestive
Diseases and Sciences 63:881--889.

\hypertarget{ref-Schmid2018}{}
13. Schmid SM, Suchodolski JS, Price JM, Tolbert MK. 2018. Omeprazole
minimally alters the fecal microbial community in six cats: A pilot
study. Frontiers in Veterinary Science 5.

\hypertarget{ref-Castellani2017}{}
14. Castellani C, Singer G, Kashofer K, Huber-Zeyringer A, Flucher C,
Kaiser M, Till H. 2017. The influence of proton pump inhibitors on the
fecal microbiome of infants with gastroesophageal refluxA prospective
longitudinal interventional study. Frontiers in Cellular and Infection
Microbiology 7.

\hypertarget{ref-Kaur2007}{}
15. Kaur S, Vaishnavi C, Prasad KK, Ray P, Kochhar R. 2007. Comparative
role of antibiotic and proton pump inhibitor in experimental
\emph{Clostridium difficile} infection in mice. Microbiology and
Immunology 51:1209--1214.

\hypertarget{ref-kaur2011effect}{}
16. Kaur S, Vaishnavi C, Prasad KK, Ray P, Kochhar R. 2011. Effect of
lactobacillus acidophilus \& epidermal growth factor on experimentally
induced \emph{Clostridium difficile} infection. The Indian journal of
medical research 133:434.

\hypertarget{ref-hung2015proton}{}
17. Hung Y-P, Ko W-C, Chou P-H, Chen Y-H, Lin H-J, Liu Y-H, Tsai H-W,
Lee J-C, Tsai P-J. 2015. Proton-pump inhibitor exposure aggravates
\emph{Clostridium difficile}--associated colitis: Evidence from a mouse
model. The Journal of infectious diseases 212:654--663.

\hypertarget{ref-Jenior2017}{}
18. Jenior ML, Leslie JL, Young VB, Schloss PD. 2017. \emph{Clostridium
difficile} colonizes alternative nutrient niches during infection across
distinct murine gut microbiomes. mSystems 2.

\hypertarget{ref-Hugenholtz2018}{}
19. Hugenholtz F, Vos WM de. 2018. Mouse models for human intestinal
microbiota research: A critical evaluation. Cellular and Molecular Life
Sciences 75:149--160.

\hypertarget{ref-Eze2017}{}
20. Eze P, Balsells E, Kyaw MH, Nair H. 2017. Risk factors for
\emph{Clostridium difficile} infections -- an overview of the evidence
base and challenges in data synthesis. Journal of Global Health 7.

\hypertarget{ref-Regrdh1985}{}
21. Regårdh C-G, Gabrielsson M, Hoffman K-J, Löfberg I, Skånberg I.
1985. Pharmacokinetics and metabolism of omeprazole in animals and man -
an overview. Scandinavian Journal of Gastroenterology 20:79--94.

\hypertarget{ref-Llorente2017}{}
22. Llorente C, Jepsen P, Inamine T, Wang L, Bluemel S, Wang HJ, Loomba
R, Bajaj JS, Schubert ML, Sikaroodi M, Gillevet PM, Xu J, Kisseleva T,
Ho SB, DePew J, Du X, Sørensen HT, Vilstrup H, Nelson KE, Brenner DA,
Fouts DE, Schnabl B. 2017. Gastric acid suppression promotes alcoholic
liver disease by inducing overgrowth of intestinal enterococcus. Nature
Communications 8.

\hypertarget{ref-Kozich2013}{}
23. Kozich JJ, Westcott SL, Baxter NT, Highlander SK, Schloss PD. 2013.
Development of a dual-index sequencing strategy and curation pipeline
for analyzing amplicon sequence data on the MiSeq illumina sequencing
platform. Applied and Environmental Microbiology 79:5112--5120.

\hypertarget{ref-Sze2019}{}
24. Sze MA, Schloss PD. 2019. The impact of DNA polymerase and number of
rounds of amplification in PCR on 16S rRNA gene sequence data. mSphere
4.

\hypertarget{ref-Schloss2009}{}
25. Schloss PD, Westcott SL, Ryabin T, Hall JR, Hartmann M, Hollister
EB, Lesniewski RA, Oakley BB, Parks DH, Robinson CJ, Sahl JW, Stres B,
Thallinger GG, Horn DJV, Weber CF. 2009. Introducing mothur:
Open-source, platform-independent, community-supported software for
describing and comparing microbial communities. Applied and
Environmental Microbiology 75:7537--7541.
